---
title: "p8105 Homework 3 (UNI:gvs2113)"
output: github_document
author: "Grace Santos"
date: "2023-10-11"
---

```{r}
library(tidyverse)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1 
This problem uses a cleaned and reduced version of a dataset entitled: "The Instacart Online Grocery Shopping Dataset 2017" that was accessed on June 24, 2017. As a leading online grocery service, Instacart allows you to shop online from local grocery stores and thus provides us with a dataset involving observations about each grocery order and its user. More specifically, the entire 2017 dataset is an anonymized dataset with over 3 million orders from over 200,000 users. With that, let's see what we can find out by using a more manageable version of the 2017 data... 

```{r}
library(p8105.datasets)
data("instacart")
```

At first glance, the `instacart` data set consists of `r ncol(instacart)` columns and `r nrow(instacart)` observations. The information provided in this data set reveals characteristics of the buyer, their buying habits, the food order and the food products included. There is data included for a total of `r n_distinct(pull(instacart, order_id))` orders placed using the Instacart app. Many of the variables in the dataset hold a collection of identification numbers for the order, the products, the customer,and the department and aisle in which the product is found. It is notable that the all of the orders in dataset we are using belong to the `train` evaluation set, indicated by the `eval_set` column. There is also informative time related data associated with the day of the week and hour of the day each order was placed. Most often, orders were placed on the  `r instacart |> pull(order_dow) |> DescTools::Mode()`th day of the week at `r instacart |> pull(order_hour_of_day) |> DescTools::Mode()`rd hour, which here means Sunday at 2pm. 

Now to answer the questions for the homework assignment: 

How many aisles are there, and which aisles are most items ordered from? 
```{r}
instacart |> 
  group_by(aisle) |> 
  summarize(count = n()) |> nrow()  # total aisles 

instacart |> 
  group_by(aisle) |> 
  summarize(count = n()) |> top_n(3) # most ordered from aisle 
```

Thus, there are 134 total aisles and the top 3 most ordered from aisles are: fresh vegetables, fresh fruits and packages vegetables fruits. 

Plot that shows the number of items ordered in each aisle: 
```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle,n)) |> 
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
This graph shows the number of items (greater than 10,000) ordered in each aisle of the grocery store for the "Instacart" data set. The lowest number of items on this graph is in the butter category and the most number of items are purchases from the fresh fruits and fresh vegetables aisles. 
 

Table showing three most popular items in specified aisles and include number of orders for each: 
```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |> 
  group_by(aisle, product_name) |> 
  summarize(items = n()) |> 
  top_n(3) |> 
  knitr::kable(digits = 1)
```

Table showing mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week: 
```{r}
instacart |> 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |> 
  group_by(product_name, order_dow) |> 
  summarize(mean_hour = mean(order_hour_of_day)) |> 
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour
  ) |> 
  knitr::kable(digits = 1)
```

## Problem 2 
```{r}
library(p8105.datasets)
data("brfss_smart2010")
```

Data Cleaning: 
```{r}
brfss_smart2010 = 
brfss_smart2010 |> 
  janitor::clean_names() |> 
  filter( 
    topic == "Overall Health",
    response %in% c("Poor", "Fair", "Good", "Very good", "Excellent")) |> 
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"), ordered = TRUE))
```

In 2002 which states were observed at 7 or more locations? 
```{r}
brfss_smart2010 |> 
  filter (year == "2002") |> 
  group_by (locationabbr) |> 
  summarize (locations = n_distinct(locationdesc)) |> 
  filter (locations >= 7) 
```
In 2002, there were 6 states that were observed at 7 or more locations.

In 2010?
```{r}
brfss_smart2010 |> 
  filter (year == "2010") |> 
  group_by (locationabbr) |> 
  summarize (locations = n_distinct(locationdesc)) |> 
  filter (locations >= 7) 
```
In 2010, there were 14 states that were observed at 7 or more locations. 

Construct a dataset limited to "Excellent" responses and then make a "spaghetti" plot of average of `data_value` across locations within a state. 
```{r}
  brfss_smart2010 |> 
  filter(response == "Excellent") |> 
  group_by(year, locationabbr, locationdesc) |> 
  summarize(avg_data_value = mean(data_value)) |> 
  ggplot(aes(x = year, y = avg_data_value, group = locationabbr, color = locationabbr)) + geom_line() + 
  labs (
    title = " Average Data Value across locations within a state for Excellent Responses",
    x = "Year",
    y = "Average Data Value",
    color = "State") 
```
The "spaghetti" plot above shows the average percent of responses classified in the "Execllent" category. The responses range from around 45% to 10% "Excellent". Overall, from 2002 to 2010, the responses stay stable at an average of around 25% "Excellent" with a slight decrease over time. 


Make a 2-panel plot showing distribution of `data_value` for responses among NY state locations 
```{r}
brfss_smart2010 |> 
  filter(locationabbr == "NY", 
         year %in% c("2006", "2010")) |> 
  ggplot(aes(x = response , y = data_value, group = locationdesc, color = locationdesc)) + geom_line() +
  facet_wrap(. ~ year) + 
  labs( 
    title = "2006 and 2010 Distribtion of Data Values by Responses among NY locations ",
    x = "Respose",
    y = "Data Value",
    color = "Location")
```
The line plot above shows the distribution of responses by locations in New York. In 2006, the majority of the responses across all locations in New York were either "Good" or "Very good". There are more locations included in the 2010 distribution and show an increase in the number of "Very good" responses with "Good" being the next most often occurring response. 


## Problem 3 
Load, tidy, merge and organize the datasets. 
```{r}
accel_data = 
  read_csv("./data/nhanes_accel.csv")

covar_data = 
  read_csv("./data/nhanes_covar.csv", skip = 4, col_names = TRUE)

mims_df = right_join(covar_data, accel_data, by = "SEQN")

mims_df_2 = 
  mims_df |> 
    filter(age >= "21") |> 
    drop_na() |> 
    mutate( 
      sex = case_match(
        sex,
        1 ~ "male", 
        2 ~ "female"),
      sex = as.factor(sex),
      education = case_match(
        education,
        1 ~ "less than HS",
        2 ~ "HS equivalent",
        3 ~ "more than HS"),
      education = factor(education, levels = c("less than HS","HS equivalent", "more than HS"), ordered = TRUE)
      ) |> 
  pivot_longer(
    min1:min1440,
    names_to = "minutes",
    names_prefix = "min",
    values_to = "MIMS")
```

Produce a reader-friendly table for the number of men and women in each education category. 
```{r}
mims_df_2 |> 
  group_by(sex, education) |> 
  summarize(n = n()) |> 
  pivot_wider(
    names_from = education,
    values_from = n) |> 
  knitr::kable(digits = 2)
```

This table shows, that there are similar numbers between sexes in participants who have either less than or more than high school education. There is a much larger difference, in favor of male participants in the category of having a high school equivalent education. The majority of participants for both sexes have more than a high school education. 

Create a visualization of age distributions for men and women in each category.
```{r}
mims_df_2 |> 
  ggplot(aes(x = education, y = age, fill = sex)) + geom_boxplot() +
  labs(
    title = "Age Distributions by Sex in each Education Category"
  )
```

From the box plots generated, you can conclude that the median ages and overall spread of ages between males and females who have less education than high school are very similar. The largest spread and difference in median are between the ages of men and women who have high school equivalent education. The one case in which the male median is higher than the female median is for those with higher than high school education. There are no ages that are greater than 80 years old included in this data. 

Aggregate across minutes to create a total activity variable for each participant over the whole day.
```{r}
mims_df_2 |> 
  group_by(SEQN, education, age, sex) |> 
  summarize(total_activity = sum(MIMS)) |> 
  ggplot(aes(x = age, y = total_activity, color = sex)) + geom_point() + geom_smooth(aes(group = sex), se = FALSE) + facet_grid(. ~ education) +
  labs(
    title = "Total Activity Variable for Each Participant by Age and Education Level"
  )
```
This plot shows the total activity of each participant across age. From the trend lines on each graph, you can conclude that for each education category, there is a decrease in total activity as age increases. Generally, women have higher amounts of total activity compared to men, except for those in the less than high school educational category. 

Make a 3 panel plot that shows 24-hour activity time courses for each education level and color by sex.  
```{r}
mims_df_2 |> 
  mutate(minutes = as.numeric(minutes)) |> 
  group_by(SEQN) |> 
  ggplot(aes(x = minutes, y = MIMS, color = sex)) + geom_line() + geom_smooth(aes(group = sex)) + facet_grid(. ~ education) + 
  labs(
    title = "24-hour activity time courses for each education level"
  )
```
This graph shows the MIMS data for each participant over a 24 hour period by education level. When looking at the trend lines for each graph, you can conclude that all education levels show a similar rounded, inverse U-shaped line. The MIMS data is generally lowest before the 500 minute mark and trends in a sharp increase until about the 600-700 minute mark. This level plateaus across the 700-1100 minute mark and then decreases until the end of the time period. This behavior is consistent with general activity related to normal sleep pattens with the most amount of activity happening in the middle, day hours of the time period. 
